Mining HIV Trends in Social Media Data

### 2016-04-14 draft/draft.tex
latest commit

\section{Introduction}

@@

We sought to determine trends in HIV and PrEP discourse on twitter, to inform and coordinate public health efforts to promote PrEP adoption and adherence for at-risk individuals. We collected 624,569 tweets related to 'HIV', 'AIDS', 'truvada', 'prophylaxis', 'imtesting', 'PrEP' from Twitter's streaming API. The tweets were restricted to English language, and the collection dates spanned from the 47th week of 2015 to the 14th week of 2016. The tweets were and cleaned of exotic characters.

@@

% show text of relevent tweets from d2v

% talk a bit about pre processing

The first analyses that we performed sought to identify certain keywords, hashtags, tweets and users, that were discussing HIV and PrEP related contexts. Word2Vec and the related method, Paragraph2Vec are unsupervised machine learning methods that have performed well at embedding natural language in a semantic vector space. In our analysis, Word2Vec allows us to determine semantically similar words, while Paragraph2Vec allows us to determine similar tweets, users and hashtags.

@@

@@

We found that topic 5 captured several keywords related to World AIDS Day (Figure 3). We can see that all of these terms peak in the 47th week of 2015 and then decline into 2016. This correlates well with the actual date of World AIDS Day, December 1st. While we aren't specifically interested in World AIDS Day to inform our understanding of PrEP discussion, this observation validates our ability to identify temporal events using DTM.

% figure 2
\begin{figure}
\centering
\includegraphics[height=2.5in, width=3.5in]{DTMfig1}
\caption{DTM topic 3 word prevalence over time. Date is YYYY-WW.}
\end{figure}

% figure 3
\begin{figure}
\centering
\includegraphics[height=2.5in, width=3.5in]{DTMfig2}
\caption{DTM topic 5 word prevalence over time. Date is YYYY-WW.}
\end{figure}

@@

Lastly we sought to create a classification scheme to classify the sentiment of tweets either positive or negative. This classifier would allow us to quickly identify positive and negative PrEP related tweets to guide public health efforts. We first performed Paragraph2Vec on both our PrEP related tweet corpus and on another tweet corpus that had binary sentiment labels, positive, or negative. Then we trained a simple logistic regression classifier on the paragraph-vectors. We found that our classifier had an accuracy of 70\% on a validation set.

We then used this classifier to classify our PrEP related tweets into positive or negative labels. We reported the most positive, and most negative tweets, by log(probability), on our full dataset, and on tweets that specifically mention either 'prep' or 'truvada' (Table 4). It seems that overall the classifier seems to capture the sentiment relatively well, however it clearly fails to recognise the sarcasm present in the most positive overall tweet.

The sentiment results aren't perfect, but they do help us identify positive individuals in PrEP and HIV public health such as user @greg0wen. The sentiment tweets also let us uncover some of the prevailing stigmas and negative sentiments surrounding PrEP usage.

% do a top positive tweets table

% do a top negative tweets table

@@

Example citation (needed right now to compile):\cite{Lamport:LaTeX}

>>>>

\section{Introduction}
Example citation (needed right now to compile):\cite{Lamport:LaTeX}

@@

We sought to determine trends in HIV and PrEP discourse on twitter, to inform and coordinate public health efforts aiming to promote PrEP adoption and adherence for at-risk individuals. We collected 624,569 tweets containing at least one of the following words 'HIV', 'AIDS', 'truvada', 'prophylaxis', 'imtesting', 'PrEP' from Twitter's streaming API. The tweets were restricted to English language, and the collection dates spanned from the 47th week of 2015 to the 14th week of 2016. The tweets were and cleaned of exotic characters.

@@

The first analyses that we performed sought to identify certain keywords, hashtags, tweets and users, that were discussing HIV and PrEP related contexts. Word2Vec and the related method, Paragraph2Vec are unsupervised machine learning methods that have performed well at embedding natural language in a semantic vector space. In our analysis, Word2Vec allows us to determine semantically similar words to a query word, while Paragraph2Vec allows us to determine similar tweets, users and hashtags to a query hashtag.

% want to do a tSNE plot, but only on things that are relatively dissimilar. For example do HIV, truvada, STDs, testing related, awareness related, ... and throw away or merge labels

@@

% might want to explain at a high level how DTM works here

@@

We found that topic 4 captured several keywords related to World AIDS Day (Figure 3). We can see that all of these terms peak in the 47th week of 2015 and then decline into 2016. This correlates well with the actual date of World AIDS Day, December 1st. Furthermore,while December 1st is World AIDS day, the whole month of Decemeber is AIDS Awarness month. We are clearly seeing the word 'awarness' peak later and last longer that the word 'WorldAIDSDay' implying that it is correlated to the whole month of December. While we aren't specifically interested in World AIDS Day, or AIDS awareness month, to inform our understanding of PrEP discussion, this observation validates our ability to identify temporal events using DTM.

% breifly mention what was in the other topics (lots of AIDS related news announcements)

% figure 2
\begin{figure}
\centering
\includegraphics[height=2.5in, width=3.5in]{DTMfig1}
\caption{DTM topic 5, word prevalence over time. Date is YYYY-WW.}
\end{figure}

% figure 3
\begin{figure}
\centering
\includegraphics[height=2.5in, width=3.5in]{DTMfig2}
\caption{DTM topic 5, word prevalence over time. Date is YYYY-WW.}
\end{figure}

@@

Lastly we sought to create a classification scheme to classify the sentiment of tweets either positive or negative. This classifier would allow us to quickly identify positive and negative PrEP related tweets to guide public health efforts. We first performed Paragraph2Vec on both our PrEP related tweet corpus and on another tweet corpus that had binary sentiment labels, positive, or negative. Then we trained a simple logistic regression classifier on the paragraph-vectors. We found that our classifier had an accuracy of 69\% on a validation set.

We then used this classifier to classify our PrEP related tweets into positive or negative labels. We reported the most positive, and most negative tweets, by log(probability), on our full dataset, and on tweets that specifically mention either 'prep' or 'truvada' (Table 4).

% do a top positive tweets table (also with tweet ID and log prob of classification)

% do a top negative tweets table (also with tweet ID and log prob of classification)

@@
